---
title: "Gambling in Victoria - Electronic Gaming Machine statistics as at 30 June 2018"
resource_files:
- SA1_SEIFA_LGA.dbf
- SA1_SEIFA_LGA.prj
- SA1_SEIFA_LGA.shp
- SA1_SEIFA_LGA.shx
runtime: shiny
output:
  flexdashboard::flex_dashboard:
    orientation: rows
    theme: flatly
    vertical_layout: scroll
    source_code: embed
---
<style type="text/css">
.navbar-brand {  /* nav_title  */
   font-size: 30px;
   font-family: Lato;
   font-weight: bold;
</style>

<style type="text/css">
.chart-title {  /* chart_title  */
   font-size: 15px;
   font-family: Lato;
</style>


```{r setup, include=FALSE}
#Load packages
library(flexdashboard)
library(rgdal)
library(leaflet)
library(dplyr)
library(magrittr)
library(readr)
library(knitr)
library(shiny)
library(stringr)
library(plotly)

```


```{r results = 'hide', message = FALSE, warning = FALSE}
#load the data

#import shp file
lga_seifa <-readOGR(dsn = ".", layer = "SA1_SEIFA_LGA")
venues_egm <- read_csv("EGM_Venues.csv")
egmStats <- read_csv("EGM_LGA_Stats.csv")
venues_egm$`Expenditure_EGM_Expenditure\n01 Jul 18 - 30 June 19` <- formatC(venues_egm$`Expenditure_EGM_Expenditure\n01 Jul 18 - 30 June 19`, format="f", big.mark=",", digits=0)
egm_seifa <- read_csv("EGM_Venues_seifa_1km_consolidated.csv")

#create icons
icon_egm <- makeIcon(
  iconUrl = "EGM_Icon-02.png")

```

Sidebar {.sidebar} 
----------------------------
**EGM expenditure and location**

<hr> </hr>

The socio-economic impact of gambling is an issue that is addressed at a state and local government level. The extent of the venue catchment influences and impacts the residents that live near it. The map and chart provides an insight into demographics surrounding these venues and highlights the potential threats of highly condense EGM venues in the most vulnerable areas. 


```{r}
#create a filter event for the choices made
hr()
selectInput("LGA", label = "Select a Local Government Area: ",
            choices = lga_seifa$LGA_NAME, selected = "Alpline (S)")

#create a reactive filter for the map base on the input
mapFilter <- reactive({
  mapF <- lga_seifa[lga_seifa$LGA_NAME == input$LGA,]
  return(mapF) })

#create a reaction filter for the venues base on the input
iconFilter <- reactive({
  icon <- venues_egm[venues_egm$LGA_NAME_2016 == input$LGA,] 
  return (icon)})

#create reactive filter for the metrics
statistics <- reactive({
  values <- filter(egmStats, LGA_NAME == input$LGA)
  return(values)})

#create reaction filter for the bubble chart based on the input
toPlot <- reactive({
  egmDensity <- filter(egm_seifa, LGA_NAME_2016 == input$LGA)
  return(egmDensity)})

```
<hr> </hr>
<font size = "1.7">***Socio-Economic Indexes for Areas (SEIFA) by Statistical Area level 1 (SA1). SEIFA scores are the State Deciles and reflects the Index of Relative Socio-Economic Advantage and Disadvantage (IRSAD)*</font>

You will notice on the map where the venues are spatially located. Using the graph you can see how many venues and EGMS that are located within the LGA and its respective SEIFA area. 
For example, if you look at Hume (C) and then Bayside (C) and compare the net expenditure, number of venues, and location of venues you will find that all 14 venues in Hume are located in disadvantaged areas, whereas in Bayside they are located in more advantaged areas but with less net spend.  

<font size="2"> Data links: 
[SEIFA Index](https://www.abs.gov.au/websitedbs/censushome.nsf/home/seifa) &nbsp;&nbsp;&nbsp;&nbsp;
[LGA Shapefile](https://discover.data.vic.gov.au/dataset/local-government-area-boundaries-property-1-25-000-vicmap-admin) <br>
[Electronic Gaming Data](https://www.vcglr.vic.gov.au/resources/data-and-research/gambling-data/gaming-expenditure-venue) <br>
Made by Nhung Seidensticker 

Row {data-width = 50}
-----------------------------------------------------------------------

### Total Adult Population {.value-box}
```{r}

renderValueBox({
  valueBox(value = prettyNum(statistics()[5], big.mark=","),caption = "Adult population",icon = "fa-users",color = "#74c476")})

```

### Total Labour force {.value-box}
```{r}
renderValueBox({
  valueBox(value = prettyNum(statistics()[9], big.mark=","), caption = "Workforce",icon = "fa-user-plus", color = "#6baed6")})

```

### Number of venues {.value-box}
```{r}
renderValueBox({
  valueBox(value = prettyNum(statistics()[2], big.mark=","),caption = "No. of venues",icon = "fa-hotel",color = "orange")})
```

### Net expenditure {.value-box}
```{r}
renderValueBox({ 
  valueBox(value = str_c("$",prettyNum(statistics()[4], big.mark=",")), 
           caption = "Net Expenditure", color = "coral")})
```

Row {data-width = 40}
-----------------------------------------------------------------------
### Total Adults per venue {.value-box}
```{r}
renderValueBox({
  valueBox(value = prettyNum(statistics()[6], big.mark=","),caption = "Adults per venue", icon = "fa-portrait", color = "#74c476")})
```

### Total Unemployed {.value-box}
```{r}
renderValueBox({
  valueBox(value = prettyNum(statistics()[10], big.mark=","),caption = "Unemployed", icon = "fa-user-minus", color = "#6baed6")})
```


### Total EGM {.value-box}
```{r}
renderValueBox({
  valueBox(value = prettyNum(statistics()[3], big.mark=","), caption = "Total EGMs",icon = "fa-dice", color = "orange")})
```

### Spend per adult {.value-box}
```{r}
renderValueBox({ 
  valueBox(value = str_c("$",prettyNum(statistics()[8], big.mark=",")), 
           icon = "fa-hand-holding-usd",caption = "Expenditure per adult", color = "coral")})
```


Row {data-height = 2500}
-----------------------------------------------------------------------
### click on a cluster to zoom in, and click on the SA1 or Pin to get more information

```{r}

renderLeaflet({

pal <- colorFactor(palette = c('#a5a5a5','#a50026','#d73027','#f46d43','#fdae61','#fee090','#e0f3f8','#abd9e9','#74add1','#4575b4','#313695'), 
                   levels = c("0","1","2","3","4","5","6","7","8","9","10"))

# pop values
SA1_popup <- paste0("<b>SA1 Code: </b>", mapFilter()$SA1_7DIG16,"<br>",
                      "<b>SEIFA Score: </b>", mapFilter()$StateDecil, "<br>",
                      "<b>SA1 Population 2016: </b>", mapFilter()$Population, "<br>",
                      "<b>Median household weekly income: </b>", 
                      str_c("$", prettyNum(mapFilter()$HHI_M_W, big.mark=",")), "<br>",
                      "<b>Median weekly rent: </b>", 
                      str_c("$", prettyNum(mapFilter()$M_Rent, big.mark=",")), "<br>",
                      "<b>Median mortgage repayment: </b>", 
                      str_c("$", prettyNum(mapFilter()$M_Mort_rep, big.mark=",")), "<br>",
                      "<b>Median weekly personal income: </b>", 
                      str_c("$", prettyNum(mapFilter()$M_WTPERI, big.mark=",")), "<br>",
                      "<b>Median Age: </b>", mapFilter()$Med_AGE)
venue_pop <- paste0("<b>Venue Name: </b>",
                    str_to_title(iconFilter()$`VENUE NAME`),"<br>",
                    "<b>Category: </b>",
                    iconFilter()$CATEGORY, "<br>",
                    "<b>Expenditure: </b>",
                    str_c("$",iconFilter()$`Expenditure_EGM_Expenditure\n01 Jul 18 - 30 June 19`))
m <- leaflet() %>%
  #setView(lng = 144.961686 , lat = -37.812791, zoom = 16) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(data = mapFilter(),
              fillColor = ~pal(StateDecil),
                  color = "grey",
                  weight = 0.5,
                  opacity = 0.6,
                  fillOpacity = 0.4,
                  popup = SA1_popup,
                  group = "SEIFA",
                  highlight = highlightOptions(weight = 5,color = "grey",bringToFront = TRUE)) %>%
  addMarkers(data = iconFilter(),
                    group = "EGM",
                    icon = icon_egm,
                    popup = venue_pop,
                   lat = ~LATITUDE, lng = ~LONGITUDE,
             clusterOptions = markerClusterOptions()) %>%
  addLegend(position = 'bottomleft',
            colors = c('#a5a5a5','#a50026','#d73027','#f46d43','#fdae61','#fee090',
                       '#e0f3f8','#abd9e9','#74add1','#4575b4','#313695'),
            labels = c("Unranked",1:10), 
            title = 'SEIFA Decile: <br>1 Disadvantaged <br> 10 Advantaged',
            opacity = 0.7) %>%
  addLayersControl(overlayGroups = c("SEIFA", "EGM"))

})

```

>Displayed is a map showing the venue location and SEIFA areas by Statistical Area level 1 (SA1). 
>The red areas and the value (1) indicates the most disadvantaged and the blue areas with value (10) indicates the most advantaged. 

### Hover over the circles to get further information on the venue

```{r}
renderPlotly({

    spendPerEGM <- toPlot()$Exependiture/toPlot()$`OPERATING EGMS`
  spendEGM <- formatC(spendPerEGM,digits=2, format = "f", big.mark = ",")
  colors <- c('#4575b4', '#74add1', '#abd9e9', '#fdae61', '#f46d43','#d73027')

p <- plot_ly(toPlot(), x = ~SEIFA_av_1km, y = ~`OPERATING EGMS`, type = 'scatter', mode = 'markers',
             color = ~Exependiture, colors = colors, size = ~Exependiture, sizes = c(10, 40),
             marker = list(opacity = 0.5, sizemode = "diameter"),
             text = ~paste("<b>Venue: </b>", str_to_title(`VENUE NAME`), 
                           "<br><b>No. EGMs: </b>", `OPERATING EGMS`,
                           "<br><b>Expenditure: ",prettyNum(str_c("$",Exependiture), 
                                                          big.mark=",", scientific = FALSE),"</b>",
                           "<br><b>per EGM: </b>", (str_c("$", spendEGM))))%>%
    colorbar(title = "Expenditure",
             outlinecolor = "transparent",
             thickness = 20) %>%
    layout(xaxis = list(title = "Median Seifa Score within 1km of EGM Venue", 
                      gridcolor = "ffffff",
                      ticklen = 5,
                      gridwidth = 1),
          yaxis = list(title = "Number of EGMs per Venue",
                      gridcolor = "ffffff",
                      ticklen = 5,
                      gridwidth = 1),
         paper_bgcolor = "#f2f2f2",
         plot_bgcolor = "#f2f2f2")
p
  
})
```

>The median seifa score was taken from the SEIFA Index for SA1 within a 1km radius of the EGM venue. The chart aims to show the number of venues within SEIFA areas followed by the number of venues. The size of the circles indicates the expenditure for the venue. 

Row {data-height = 50}
-----------------------------------------------------------------------
### References

<font size="1.5">
Victorian Commission for Gambling and Liquor Regulation. (2019).Current LGA population density and gaming expenditure statistics [Data file]. Retrieved from https://www.vcglr.vic.gov.au/resources/data-and-research/gambling-data/population-density-and-gaming-expenditure
<br>
Victorian Commission for Gambling and Liquor Regulation. (2019). Current gaming expenditure by venue [Data file]. Retrieved from https://www.vcglr.vic.gov.au/resources/data-and-research/gambling-data/gaming-expenditure-venue. <br>
Australian Bureau of Statistics. (2018). Local Government Area, Indexes, SEIFA 2016 [Data file]. Retrieved from https://www.abs.gov.au/AUSSTATS/abs@.nsf/DetailsPage/2033.0.55.0012016?OpenDocument. <br>
Department of Environment, Land, Water & Planning. (2019). Local Government Area Boundaries (Property) 1:25,000 - Vicmap Admin [Data file]. Retrieved from https://discover.data.vic.gov.au/dataset/local-government-area-boundaries-property-1-25-000-vicmap-admin. <br>
Australian Bureau of Statistics. (2016). Victoria Local Government Area ASGS Edition 2016 in .csv Format [Data file]. Retrieved from https://www.abs.gov.au/AUSSTATS/abs@.nsf/DetailsPage/1270.0.55.003July%202016?OpenDocument
